Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeast Taiwan

### 2019-10-09 analysis/paper/paper.Rmd
added sentences to method section

# pottery shape analysis 
- review about pottery shape analysis recently

@@

## geometric morphometric approach

The drawings of pottery with half cross-section from rim to body were imported into Inkscape software for outline digitalization. All half cross-section images were copied and flipped to create enclosed images in order to be ran in the R environment. The digitised outlines were converted into a list of x-y pixel coordinates for elliptic Fourier analysis (EFA), which can assess pottery morphological difference among pottery outlines from three occupation contexts. The EFA generates harmonic coefficients were analysed by principal component analysis (PCA) to identity major patterns in data through dimensionality reduction. The principal components (PCs) were analysed with a multivariate analysis of variance (MANOVA) to test for significant effects of the occupation context on shape variances. Finally, we computed a significance test for the equality of coefficients of variation (CVs) among multiple groups of PCs to understand the differences in shape standardisation. Outline analyses were carried out using the package Momocs in the software R.

>>>>

# Shape analysis for artefacts
- review about pottery shape analysis recently

There are two different approaches for shape analysis, one is landmark analysis and another one is outline analysis. In this paper, we conducted outline analysis because it is useful to compare geometric structures where individual homologous landmarks are difficult to locate.

GMM analysis: measuring standardisation through shape

@@

## outline analysis approach

The scanned cross-section drawings for pots covering from rim to bottom were acquired from Bureau of Cultural Affairs in Yilan. Scanned drawings were imported into Inkscape software for outline tracing to remove some additional information on the orginal drawings. Each finished half cross-section image was duplicated, flipped, and then joined with original one to create a complete cross-section image in order to be performed in the R software (www.rproject.org, Core-Team, 2015). Geometric morphometric analyses were conducted in R using the functions included in the Momocs, a package intended to quantify the shape and compare its variation, especially for outline analysis [@Bonhomme2014]. The digitised outlines were converted into a list of x-y pixel coordinates for elliptic Fourier analysis (EFA), which assesses morphological differences among pottery shapes from three occupation contexts. EFA is powerful method to extract geometric information and decompose the shape into a series of trigonometric functions referred as harmonic coefficients or harmonics [@Claude2008; @Bonhomme2014]. Harmonic..
The EFA generates harmonic coefficients were analysed by principal component analysis (PCA) to identity major patterns of variation among the shape data through dimensionality reduction. The principal components (PCs) were analysed with a multivariate analysis of variance (MANOVA) to test for significant effects of the occupation context on shape variances. Finally, we computed a significance test to assess the equality of coefficients of variation (CVs) among multiple groups of PCs to understand the differences in shape standardisation. The R package cvequality (Version 0.2.0; Marwick and Krishnamoorthy 2019) in the CRAN repository was used to conduct the significance test.
